#Exploratory Analysis of McDonalds Menu
##Galina Endarova

##Overview
This markdown explores a McDonalds menu, trying to find some correlations 
between the dietary properties and see if there is a safe choice in there.

##Data
The data has been downloaded from Kaggle.

##Analysis

```{r, fig.align = 'center', fig.width = 12}
#reading in the data
knitr::opts_chunk$set(fig.width = 11)
suppressWarnings(suppressPackageStartupMessages(library(data.table)))
menu <- data.table(read.csv("Data/menu.csv", stringsAsFactors = FALSE))
corrData <- menu[,-c(1,2,3,5,7,9,12,14,16,18)]
corrplot::corrplot(cor(corrData), method = "number")
```


From the Correlations plot (Appendix) we can see that there is a strong 
relationship between calories and each of the macro groups - Fat, Protein and
Carbohydrates. That makes sense - those calories have to come from somewhere. 

Cholesterol and Calories are positively correlated. So are Calories and Sodium 
and Cholesterol and Sodium.

###Ratio
We are probably not going to find the perfect meal in McDonalds. However, 
it is worth exploring whether a decent Carbs-Protein-Fat ratio can be found 
in a single item on the menu.
```{r}
menu[, totalGrams := Total.Fat + menu$Protein + Carbohydrates]
menu[, CarbsRatio := Carbohydrates / totalGrams]
menu[, FatRatio := Total.Fat / totalGrams]
slim <- nrow(menu[CarbsRatio < 0.6 & CarbsRatio > 0.4 & FatRatio < 0.25
          & FatRatio > 0.15, ])
```
With a slim check about a balance of fat and carbs we see that 
`r slim` / `r slim/nrow(menu)*100`% out of the items on the menu follow that 
ratio. The small number is not that surprising - we would expect McDonalds items to be full of
fat and carbs (hello, french fries!).

###The biggest health risks 
The biggest health risks are probably the items with the most 
calories and where most calories come from either fat or carbs?
Not entirely sure about this one -> fat can be very satiating. However,
if you get a calorie bomb loaded with carbs, you could be crashing two hours 
later.


The biggest calorie bomb is not that surprising - 
`r menu[Calories == max(Calories), Item]` have 
`r menu[Calories == max(Calories), Calories]` calories. Perhaps most people
do not eat 40 nuggets on their own? Sharing is caring.

```{r}
setorder(menu, Calories)
DT::datatable(tail(menu[, c("Item", "Calories")]))
```

It is a biut of a comfort that the highest food by calories is far removed from 
the rest of them. However, it is a bit surprising that the next 4 largest meals 
are different breakfasts, given that what I have observed are bagels and 
english muffins, not a hearty spread.


A deeper look into the relationship between sugar and fiber is interesting. 
Sugars go all the way to up to 128 grams, while Fiber is all the way down at 
6 or less grams. As a reminder, the 
[sugar recommendation](https://www.nhs.uk/live-well/eat-well/how-does-sugar-in-our-diet-affect-our-health/) 
for adults is no more than 30g daily. At the same time, the 
[recommended fiber intake](https://www.nhs.uk/live-well/eat-well/how-to-get-more-fibre-into-your-diet/)
is at 30g per day as well.
And yet you would be hard pressed to obtain anything close to the daily 
recommended value using a McDonalds diet. If we try to eat following the sugar 
recommendation, we would be able to eat 
`r nrow(menu[Sugars < 30]) / nrow(menu)`% of the items. So not so bad.

```{r}
# TODO: need to make the points bigger
suppressWarnings(library(ggplot2))
ggplot(menu, aes(x = Sugars, y = Dietary.Fiber, col = Category)) +
  geom_point()
```


So with the exception of the chicken nuggets, it seems like the 
shakes and smoothies are the highest offenders in terms of calories and sugars.
```{r fig.width=12}
# TODO: enlarge the plot
ggplot(menu, aes(x = Category, y = Calories)) +
  geom_boxplot()
```

In terms of sugar, it is just as unsurprising that all the beverages contain
high amounts of sugar. It is a bit telling that the mediums of all
the beverages are higher than that of the desserts.
```{r}
ggplot(menu, aes(x = Category, y = Sugars)) +
  geom_boxplot()
```

Highest in fiber foods are, unsurprisingly, the salads. A bit depressing
that when you go to McDonalds the best thing to eat is the same thing you would 
be eating everywhere else.

```{r}
ggplot(menu, aes(x = Category, y = Dietary.Fiber)) +
  geom_boxplot()
```

To drive the point home about the daily recommended values,  
```{r}
menu[, DailyPropSugar := Sugars / 30]
menu[, DailyPropFiber := Dietary.Fiber / 30]
menu[, DailyPropCalories := Calories / 2000]
```

Plot the proportion of sugars as part of the daily recommended intake as a 
boxplot again. - 30g of sugar for adults

The smoothies and other beverages are understandable - I was prepared for 3 time the recommended daily intake. However, Jesus about the salads! 
Why does a salad have any sugars? Unless they count fruit salads as well?
```{r}
ggplot(menu, aes(x = Category, y = DailyPropSugar)) +
  geom_boxplot() + 
  ggtitle("Proportion of the daily sugar limits by category")
```

Southwest Salad has a lot of sugar - 15 grams. How come? Dressing?
```{r}
menu[Category == "Salads", Item]
```

```{r}
# TODO: make the scales the same
ggplot(menu, aes(x = DailyPropCalories, y = DailyPropFiber)) +
  geom_point()
```

Calculate fiber as a percentage of total grams? Or fiber to fat ratio?
Fiber to sugar ratio?

Let's check the correlations again.
```{r}
cor_names <- names(menu)[
  !(grepl("Daily", names(menu))) &
    !(grepl("Prop", names(menu))) &
    !(grepl("Ratio", names(menu))) &
    !(names(menu) %in% c("Category", "Item", "Serving.Size"))]

corrplot::corrplot(cor(menu[,.SD, .SDcols = c(cor_names)]), method = "number")
```

Carbohydrates and sodium don't seem to go together. But Sodium and 
(Fat + Protein) do, as well as Sugars and Carbohydrates. 

Given that sugars are a type of carbohydrate, that makes lots of sense.
On the other hand, something has to give flavour to the other items, 
and that is typically fat and salt.