#Exploratory Analysis of McDonalds Menu
##Galina Endarova

##Overview
This markdown explores a McDonalds menu, trying to find some correlations 
between the dietary properties and see if there is a safe choice in there.

##Data
The data has been downloaded from Kaggle.

##Analysis

```{r, fig.align = 'center', fig.height = 9}
#reading in the data
suppressWarnings(suppressPackageStartupMessages(library(data.table)))
menu <- data.table(read.csv("Data/menu.csv", stringsAsFactors = FALSE))
corrData <- menu[,-c(1,2,3,5,7,9,12,14,16,18)]
corrplot::corrplot(cor(corrData), method = "number")
```


From the Correlations plot (Appendix) we can see that there is a strong 
relationship between calories and each of the macro groups - Fat, Protein and
Carbohydrates. That makes sense - those calories have to come from somewhere. 

Cholesterol and Calories are positively correlated. So are Calories and Sodium 
and Cholesterol and Sodium.

###Ratio
We are probably not going to find the perfect meal in McDonalds. However, 
it is worth exploring whether a decent Carbs-Protein-Fat ratio can be found 
in a single item on the menu.
```{r}
menu[, totalGrams := Total.Fat + menu$Protein + Carbohydrates]
menu[, CarbsRatio := Carbohydrates / totalGrams]
menu[, FatRatio := Total.Fat / totalGrams]
slim <- nrow(menu[CarbsRatio < 0.6 & CarbsRatio > 0.4 & FatRatio < 0.25
          & FatRatio > 0.15, ])
```
With a slim check about a balance of fat and carbs we see that 
`r slim` / `r slim/nrow(menu)*100`% out of the items on the menu follow that 
ratio. So not that surprising - we would expect McDonalds items to be full of
fat and carbs (hello, french fries!).

###The biggest health risks 
The biggest health risks are probably the items with the most 
calories and where most calories come from either fat or carbs?
Not entirely sure about this one -> fat can be very satiating. However,
if you get a calorie bomb loaded with carbs, you could be crashing two hours 
later.
```{r}
cat(menu[Calories == max(Calories), Item], " have ", 
        menu[Calories == max(Calories), Calories], " calories.")
```

The biggest calorie bomb is not that surprising, but perhaps most people
do not eat 40 nuggets on their own? Sharing is caring.

```{r}
setorder(menu, Calories)
DT::datatable(tail(menu[, c("Item", "Calories")]))
```

Well, they were not kidding about the highest amount of calories.
A surpise for me here was that the next 4 biggest calorie bombs were different
breakfasts, given that what I have observed are bagels and english muffins, 
not a hearty spread.


Sugars go all the way to up to 128 grams, while Fiber is all the way down at 
6 or less grams. I need to have a look at what the difference is between 
Fiber and Fiber daily value. 

```{r}
# TODO: need to make the points bigger
suppressWarnings(library(ggplot2))
ggplot(menu, aes(x = Sugars, y = Dietary.Fiber, col = Category)) +
  geom_point()
```

It would be interesting to create a boxplot of calories, fiber and so on per 
category. Or just in general.

So with the exception of the chicken nuggets, it seems like the 
shakes and smoothies are the highest offenders in terms of calories.
```{r}
# TODO: enlarge the plot
ggplot(menu, aes(x = Category, y = Calories)) +
  geom_boxplot()
```

In terms of sugar, it is just as unsurprising that all the beverages contain
high amounts of sugar. It is a bit telling that the mediums of all
the beverages are higher than that of the desserts.
```{r}
ggplot(menu, aes(x = Category, y = Sugars)) +
  geom_boxplot()
```

Highest in fiber foods are, unsurprisingly, the salads. A bit depressing
that when you go to McDonalds the best thing to eat are salads. It is 
completely reasonable, but I just don't like it. 

```{r}
ggplot(menu, aes(x = Category, y = Dietary.Fiber)) +
  geom_boxplot()
```

```{r}
menu[, DailyPropSugar := Sugars / 30]
menu[, DailyPropFiber := Dietary.Fiber / 30]
menu[, DailyPropCalories := Calories / 2000]
```

Plot the proportion of sugars as part of the daily recommended intake as a 
boxplot again. - 30g of sugar for adults

The smoothies and other beverages are understandable - I was prepared for 3 time the recommended daily intake. However, Jesus about the salads! 
Why does a salad have any sugars? Unless they count fruit salads as well?
```{r}
ggplot(menu, aes(x = Category, y = DailyPropSugar)) +
  geom_boxplot()
```

Southwest Salad has a lot of sugar - 15 grams. How come? Dressing?
```{r}
menu[Category == "Salads", Item]
```

Plot the proportion of sugars as daily recommended agains the proportion of 
fiber as daily recommended. - 30g 
Plot fiber against calories. 
```{r}
# TODO: make the scales the same
ggplot(menu, aes(x = DailyPropCalories, y = DailyPropFiber)) +
  geom_point()
```

Calculate fiber as a percentage of total grams? Or fiber to fat ratio?
Fiber to sugar ratio?

Let's check the correlations again.
```{r}
cor_names <- names(menu)[
  !(grepl("Daily", names(menu))) &
    !(grepl("Prop", names(menu))) &
    !(grepl("Ratio", names(menu))) &
    !(names(menu) %in% c("Category", "Item", "Serving.Size"))]

corrplot::corrplot(cor(menu[,.SD, .SDcols = c(cor_names)]), method = "number")
```

Carbohydrates and sodium don't seem to go together. But Sodium and 
(Fat + Protein) do, as well as Sugars and Carbohydrates. 

Given that sugars are a type of carbohydrate, that makes lots of sense.
On the other hand, something has to give flavour to the other items, 
and that is typically fat and salt.