#Exploratory Analysis of McDonalds Menu
##Galina Endarova

##Overview
##Analysis

```{r echo=FALSE}
#reading in the data
menu<-read.csv("Data/menu.csv", stringsAsFactors = FALSE )
str(menu)
menu$Item[which(menu$Calories==max(menu$Calories))]
DT::datatable(menu)
corrData<-menu[,-c(1,2,3,5,7,9,12,14,16,18)]
corrplot::corrplot(cor(corrData), method = "number")
sapply(corrData, class)
names(corrData)
```
From the Correlations plot (Appendix) we can see that there is a strong relationship between calories and each of the macro groups - Fat, Protein and Carbohydrates. That makes sense - those calories have to come from somewhere. 

Cholesterol and Calories are positively correlated. So are Calories and Sodium and Cholesterol and Sodium.

###Ratio
We are probably not going to find the perfect meal in McDonalds. However, it is worth exploring whether a decent Carbs-Protein-Fat ratio can be found in a single item on the menu. 

```{r}
menu$totalGrams<-menu$Total.Fat + menu$Protein + menu$Carbohydrates
menu$CarbsRatio<-menu$Carbohydrates/menu$totalGrams
menu$FatRatio<-menu$Total.Fat/menu$totalGrams
nrow(menu[menu$CarbsRatio<0.6 & menu$CarbsRatio>0.4 
          & menu$FatRatio<0.25 & menu$FatRatio>0.15,])

```

###The biggest health risks 

```{r}
menu$Item[which(menu$Calories==max(menu$Calories))]
```
